---
title: "Project1 Presentation"
author: "Yixiao Dong"
date: "10/7/2019"
output: html_document
---
# Background
UMD(Urban Ministries of Durham) is a non-profit orgnization which connects with the community to end homelessness and fight poverty. This project is designed to dig into the data provided by UMD and find something valuable for the staff who are working for the orgnization. 

# Raw Data
The data is provided by UMD which recorded the aids they give to people in need in the past years. The raw data sources file("UMD_Services_Provided_20190719.tsv") is given in the data folder. The file mainly includes the following information

* Date  
* Client File Number: Client identifier (individual or family)  
* Client File Merge: Separate files created for one family are merged together  
* Bus Tickets: The service discontinued  
* Notes of Service  
* Food Provided for: Number of people in the family for which food was provided  
* Food Pounds  
* Clothing Items  
* Diapers  
* School Kits  
* Hygiene Kits  
* Financial Support: Money provided to clients. Service discontinued.

# Goal
In this project, imagine when a client come for help, I want to predict whether he will come again based on the aid he get in this visit and the imformation about his previous visits (if he visited before). In the way, we can pay more attention to the people who is likely to come again because they may need more help.

# Procedure
* data exploration  
* data process
* classification model

# Data exploration
#### Time
```{R include=FALSE}
library(tidyverse)

#read data
df <- read_tsv("/Users/Yixiao/Desktop/611project/project_1/data/UMD_Services_Provided_20190719.tsv")

#remove useless columns
df <- select(df, -`Notes of Service`, -`Type of Bill Paid`, -`Payer of Support`, -`Payer of Support`, -`Referrals`, -`Field1`, -`Field2`, -`Field3`)

#change date type from mon/day/year (str) to year, month, day (integer)
df <-separate(df, Date, into = c("month", "day", "year"), sep = "/")
df <- df[,c(3,1,2,4:13)]
df$month <- as.numeric(df$month)
df$day <- as.numeric(df$day)
df$year <- as.numeric(df$year)
```
```{r echo=FALSE}
#count numbers of records in each year
count_by_year <- df %>% group_by(year) %>% summarise(count = n())
ggplot(count_by_year) +
  geom_point(aes(year, count), size=1.5) +
  ylab("Number of Records") +
  labs(title=("Numbers of Records in Each Year"))
```
```{r echo=FALSE}
#count numbers of records in each month in year 1999 to 2001 and 2017 to 2019
count_by_month <- df %>% 
  filter((year>=2000 & year<=2002)|(year>=2017 & year <=2019)) %>% 
           group_by(year, month) %>% 
           summarise(count = n())
ggplot(count_by_month) +
  geom_bar(aes(month, count), stat="identity") +
  facet_wrap(~year, nrow=2) +
  ylab("Number of Records") +
  scale_x_continuous(breaks=seq(3, 12, 3)) +
  labs(title=("Numbers of Records in Each Month in 1999 to 2001 and 2017 to 2019"))
```
```{r echo=FALSE}
df <- df %>% filter((year>2001 & year <2019)) %>% arrange(year,month,day)
```
#### Type of Service
```{r echo=FALSE}
#count number of records of different type of service in each year (remove records with 0 and NA)
type_df <- df %>% gather(`Bus Tickets (Number of)`, `Food Provided for`,`Food Pounds`, `Clothing Items`, `Diapers`, `School Kits`, `Hygiene Kits`, `Financial Support`, key="Type", value="number") %>% filter(number!=0 & is.na(number)==FALSE)
count_by_type <- type_df %>% group_by(year, Type) %>% summarise(count = n())
ggplot(count_by_type) +
  geom_bar(aes(year, count), stat="identity") +
  facet_wrap(~Type, nrow=2) +
  ylab("Number of Records") +
  labs(title=("Number of Records of Different Type of Service in Each Year"))
```
```{r echo=FALSE}
#Food Pounds record in 2005 and 2006
food_pounds <- df %>% 
  filter((year == 2006 | year == 2005)& `Food Pounds` != 0 & is.na(`Food Pounds`) == FALSE) %>% 
  select(month,year, `Food Pounds`) %>%
  group_by(year, month) %>% 
  summarise(count = n())
ggplot(food_pounds) +
  geom_bar(aes(month, count), stat="identity") +
  facet_wrap(~year) +
  scale_x_continuous(breaks=seq(0, 12, 3)) +
  labs(title=("Food Pounds Records in 2005 and 2006"))
```

# Data Processing

1. Select data from 2005.12 to 2019.7
2. Select "Food Provided for", "Food Pounds", "Clothing Items" 
3. Fill Na with 0 
4. Merge "Client file Number" and "Client file Merge" 
5. Add "Come Again": 0 means not visit again, 1 means visit again 
6. Compute and add: "Number of Visits before", "Days Since Last Visit" 
7. Compure and add: "Avg Food Provided For Before", "Avg Food Pounds Before", "Avg Clothing Items Before" 
8. Save the dataset.

# Classification Model
In this project, I use XGBoost Model.
```{r include=FALSE}
library(tidyverse)
library(caret)
require(xgboost)
```
#### read data
The data used in prediction model includes following columns:

* Come Again
* Food Provided For
* Food Pounds
* Clothing Items
* Number of Visits Before
* Days Since last Visit
* Avg Food Provided For Before
* Avg Food Pounds Before
* Avg Clothing Items Before

```{r}
df <- read.csv("/Users/Yixiao/Desktop/611project/project_1/data/data.csv")
df$Come.Again <- as.factor(df$Come.Again)
df <- select(df, -year, -month, -day, -`Client.File.Number`)
df <- df[,c(4,1:3,5:9)]
```
#### split train data and test data
```{r}
set.seed(0)
train_i <- createDataPartition(df$Food.Provided.for, p=0.8, list=FALSE)
train_df <- df[train_i,]
test_df <- df[-train_i,]
```

#### train data is unbalanced
```{r}
count_train_df <- train_df %>% group_by(Come.Again) %>% summarise(count=n())
ggplot(count_train_df)+
  geom_bar(aes(Come.Again,count), stat = "identity") +
    labs(title = "Train Data is Unbalanced")
```

#### reconstruct train data by over-sampling
```{r}
train_df0 <- filter(train_df, Come.Again==0)
train_df1 <- filter(train_df, Come.Again==1)
train_df <- rbind(train_df0, train_df0, train_df0, train_df0, train_df1)
```

#### split train and test data into label and features
```{r}
train_label <- train_df %>% select(`Come.Again`) %>% as.matrix()
train_feature <- train_df %>% select(-`Come.Again`) %>% as.matrix()
test_label <- test_df %>% select(`Come.Again`) %>% as.matrix()
test_feature <- test_df %>% select(-`Come.Again`) %>% as.matrix()
```

#### train xgboost model
```{r results="hide"}
xgb <- xgboost(data = train_feature, 
               label = train_label, 
               max_depth = 5,
               nround = 2000,
               subsample = 0.8,
               colsample_bytree = 0.5,
               eval_metric = "logloss",
               objective = "binary:logistic"
)
```

#### predict function returns a probablity in (0,1)
```{r}
xgb.probs = predict(xgb , train_feature)
hist(xgb.probs)
```

####  select probabilty >0.6 as Come again, <0.6 as not Come again
```{r}
test_prob <- predict(xgb, test_feature)
test_pred <- test_prob >0.6
```

#### evaluate results
```{r}
tp = sum(test_label==1&test_pred==1)
fp = sum(test_label==0&test_pred==1)
tn = sum(test_label==0&test_pred==0)
fn = sum(test_label==1&test_pred==0)
accuracy = (tp+tn)/(tp+fp+tn+fn)
true_positive_rate = tp/(tp+fn)
true_negtive_rate = tn/(tn+fp)
```
For records that people come again, there are 10790 in total, we predict 7666 correctly. The true positive rate is 0.71.  
For records that people didn't come again, there are 2668 in total, we predict 1710 correctly. The ture negtive rate is 0.64.  
For all the records, the accuracy is 0.70.

#### plot feature importance
```{r  results="hide"}
imp_matrix <- xgb.importance(feature_names = colnames(train_feature), model = xgb)
print(xgb.plot.importance(importance_matrix = imp_matrix))
```

#### relationship between number of visits before and come again
```{r}
df1 <- df %>% 
  filter(`Number.of.Visits.before`<50) %>% 
  group_by(`Number.of.Visits.before`) %>% 
  summarise(Come.Again.rate = sum(`Come.Again`==1)/n())
ggplot(df1)+
  geom_line(aes(`Number.of.Visits.before`,`Come.Again.rate`))
```

# Conclusion
1. Data from 2001.7 to 2019.7 is more valid.
2. Most people come for cloth and food. And the food pounds record actually starts at about 2005.12.
3. By using XGBoost, we can accurately predict whether a client will come again based on the aid he get in this visit and the imformation about his previous visits. (we can give a probability) Accuracy:0.70, True Positive Rate:0.71, True Negtive Rate:0.64.
4. Numbers of visits before contributes the most to whether a client will come again. The more number of visits before, the more likely he is to come again.
